---
title: "Video Game Sales — Preprocessing (Through Missing Value Handling)"
author: "Your Name"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width  = 10,
  fig.height = 5,
  dpi = 150
)
```

## Setup

```{r libs}
library(tidyverse)
library(scales)
library(knitr)

source(here::here("R", "01_load_clean.R"))
```

---

## 1. Load Raw Data

```{r load}
raw <- load_raw()
glimpse(raw)
```

---

## 2. Missing Value Overview

```{r missing-table}
col_summary(raw) %>%
  arrange(desc(null_pct)) %>%
  kable(digits = 2, caption = "Missing Values — Raw Dataset")
```

```{r missing-plot, fig.height=4}
col_summary(raw) %>%
  filter(null_pct > 0) %>%
  ggplot(aes(x = reorder(column, null_pct), y = null_pct)) +
  geom_col(fill = "steelblue", color = "white") +
  coord_flip() +
  labs(title = "Missing Value Rate by Column",
       x = NULL, y = "Missing (%)") +
  theme_minimal()
```

> **Observations:**
> - `Critic_Score`, `User_Score`, `Critic_Count`, `User_Count`, `Developer`, and `Rating` all have substantial missingness — these come from Metacritic and were not available for every title.
> - `Year_of_Release` and `Genre` have minor missingness.
> - Sales columns and `Name`/`Platform`/`Publisher` are nearly complete.

---

## 3. Cleaning Steps

### 3a. Remove Missing / Zero Target

We cannot model rows with no `Global_Sales` value.

```{r clean-target}
df <- raw %>% clean_target()
cat("Rows remaining:", nrow(df), "\n")
```

### 3b. Clean Year of Release

`Year_of_Release` is stored as character (some entries are `"N/A"`). We coerce to integer and drop implausible values outside 1980–2020.

```{r clean-year}
df <- df %>% clean_year()
cat("Year range:", range(df$Year_of_Release), "\n")
```

### 3c. Filter Valid Genres

Drop rows with missing or unrecognised genre values.

```{r clean-genre}
df <- df %>% clean_genre()
cat("Genres present:", paste(sort(unique(df$Genre)), collapse = ", "), "\n")
```

### 3d. Clean User Score

`User_Score` contains `"tbd"` strings where a score hasn't been assigned yet. We replace these with `NA` and convert the column to numeric.

```{r clean-user-score}
df <- df %>% clean_user_score()

cat("User_Score — NA count after cleaning:",
    sum(is.na(df$User_Score)), "\n")
```

### 3e. Standardise ESRB Rating

Non-standard ESRB values are collapsed to `"Other"` and missing ratings are labeled `"Unknown"`.

```{r clean-rating}
df <- df %>% clean_rating()
df %>% count(Rating, sort = TRUE) %>% kable()
```

---

## 4. Cleaned Dataset Summary

```{r cleaned-summary}
cat("Shape after cleaning:", nrow(df), "rows x", ncol(df), "cols\n\n")

col_summary(df) %>%
  arrange(desc(null_pct)) %>%
  kable(digits = 2, caption = "Missing Values — After Cleaning")
```

> **Remaining missingness** in `Critic_Score`, `User_Score`, `Rating`, etc. reflects games that simply were not reviewed on Metacritic or rated by the ESRB — this is expected and will be addressed at the modeling stage.

---

## 5. Stopping Point

The dataset is now clean through the missing value handling stage. Remaining steps to complete:

- **Feature engineering** — log-transform of `Global_Sales`, scaling `User_Score` to 0–100, decade variable, platform grouping
- **EDA** — distributions, genre/platform trends, score vs sales relationships
- **Regression modeling** — OLS with `Critic_Score`, `User_Score`, `Genre`, `Rating`, `Year_of_Release`

```{r save}
write_csv(df, here::here("data", "vgsales_clean.csv"))
cat("Saved: data/vgsales_clean.csv\n")
```
