---
title: "Video Game Sales Final Model Selection & Analysis"
author: "Dangelo Feliciano, Anirudh Eswara"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 5,
  dpi = 150
)
```

```{r libs}
library(tidyverse)
library(knitr)
library(lmtest)
library(car)

source(here::here("R", "01_load_clean.R"))
```

## Data Setup

```{r prep}
df <- run_pipeline()

df <- df %>%
  mutate(
    log_Global_Sales = log1p(Global_Sales),
    User_Score_100 = User_Score * 10,
    Genre = as.factor(Genre),
    Rating = if_else(Rating %in% c("E", "E10+", "T", "M"), as.character(Rating), "Other"),
    Rating = factor(Rating, levels = c("E", "E10+", "T", "M", "Other"))
  )

model_df <- df %>%
  filter(
    !is.na(Critic_Score),
    !is.na(User_Score_100),
    !is.na(Rating)
  )

set.seed(42)
folds <- sample(rep(1:10, length.out = nrow(model_df)))

cat("rows:", nrow(model_df), "\n")
cat("Genre baseline:", levels(model_df$Genre)[1], "\n")
cat("Rating baseline:", levels(model_df$Rating)[1], "\n")
```

## Getting to the Final Model

### Full Model

We started with everything we thought mattered.

```{r full_model}
full_model <- lm(
  log_Global_Sales ~ Critic_Score + User_Score_100 + Year_of_Release + Genre + Rating,
  data = model_df
)

summary(full_model)
```

```{r cv_full}
rmse_cv <- function(formula, data, folds) {
  rmse <- numeric(10)
  for (k in 1:10) {
    train <- data[folds != k, ]
    test <- data[folds == k, ]
    fit <- lm(formula, data = train)
    pred <- predict(fit, newdata = test)
    rmse[k] <- sqrt(mean((test$log_Global_Sales - pred)^2))
  }
  mean(rmse)
}

rmse_full <- rmse_cv(
  log_Global_Sales ~ Critic_Score + User_Score_100 + Year_of_Release + Genre + Rating,
  model_df, folds
)
cat("Baseline CV RMSE:", round(rmse_full, 4), "\n")
```

### Stepwise - No Improvement

Stepwise selection didn't drop anything, so all predictors are earning their place.

```{r stepwise, results='hide'}
step_model <- step(full_model, direction = "both", trace = 0)
```

```{r stepwise_results}
cat("Full AIC:", round(AIC(full_model), 2), "\n")
cat("Step AIC:", round(AIC(step_model), 2), "\n")
cat("Dropped:", setdiff(names(coef(full_model)), names(coef(step_model))), "\n")
```

### Critic Score x User Score - Best Model

We wanted to know whether games that are well received by both critics and users do even better than you'd expect from either score alone. This turned out to be our best model, it had the lowest AIC and CV RMSE of everything we tried. The ANOVA below confirms the interaction term is a statistically significant improvement.

```{r int_scores}
model_int_scores <- lm(
  log_Global_Sales ~ Critic_Score * User_Score_100 + Year_of_Release + Genre + Rating,
  data = model_df
)

summary(model_int_scores)

rmse_int_scores <- rmse_cv(
  log_Global_Sales ~ Critic_Score * User_Score_100 + Year_of_Release + Genre + Rating,
  model_df, folds
)
cat("CV RMSE:", round(rmse_int_scores, 4), "\n")
cat("AIC:", round(AIC(model_int_scores), 2), "\n")
```

```{r int_scores_anova}
anova(full_model, model_int_scores)
```

### Polynomial Year - Didn't Beat the Interaction Model

We thought year might curve since the industry peaked around 2008. The quadratic term helped over the baseline but couldn't match the score interaction.

```{r poly_year}
model_poly <- lm(
  log_Global_Sales ~ Critic_Score + User_Score_100 + poly(Year_of_Release, 2) + Genre + Rating,
  data = model_df
)

summary(model_poly)

rmse_poly <- rmse_cv(
  log_Global_Sales ~ Critic_Score + User_Score_100 + poly(Year_of_Release, 2) + Genre + Rating,
  model_df, folds
)
cat("CV RMSE:", round(rmse_poly, 4), "\n")
cat("AIC:", round(AIC(model_poly), 2), "\n")
```

```{r poly_anova}
anova(full_model, model_poly)
```

### Critic Score x Genre - Didn't Beat the Final Model

This one was interesting because it gets at our research question directly does critical reception matter more in some genres than others? The answer seems to be yes based on the F-test, but the model still had a higher AIC and CV RMSE than our Critic x User Score model so we didn't pick it as the final.

```{r int_genre}
model_int_genre <- lm(
  log_Global_Sales ~ Critic_Score * Genre + User_Score_100 + poly(Year_of_Release, 2) + Rating,
  data = model_df
)

summary(model_int_genre)

rmse_int_genre <- rmse_cv(
  log_Global_Sales ~ Critic_Score * Genre + User_Score_100 + poly(Year_of_Release, 2) + Rating,
  model_df, folds
)
cat("CV RMSE:", round(rmse_int_genre, 4), "\n")
cat("AIC:", round(AIC(model_int_genre), 2), "\n")
```

```{r int_genre_anova}
anova(model_poly, model_int_genre)
```

## Model Comparison

```{r comparison}
comparison <- data.frame(
  Model = c(
    "Full Model",
    "Stepwise AIC",
    "Critic x User Score (Final)",
    "Polynomial Year",
    "Critic x Genre + Poly Year"
  ),
  AIC = round(c(
    AIC(full_model), AIC(step_model), AIC(model_int_scores),
    AIC(model_poly), AIC(model_int_genre)
  ), 2),
  CV_RMSE = round(c(
    rmse_full, rmse_cv(formula(step_model), model_df, folds),
    rmse_int_scores, rmse_poly, rmse_int_genre
  ), 4),
  Adj_R2 = round(c(
    summary(full_model)$adj.r.squared, summary(step_model)$adj.r.squared,
    summary(model_int_scores)$adj.r.squared, summary(model_poly)$adj.r.squared,
    summary(model_int_genre)$adj.r.squared
  ), 4)
)

kable(comparison, caption = "Model Comparison")
```

The Critic x User Score interaction model wins on both AIC (7276) and CV RMSE (0.4098). We're going with this one.

## Final Model Deep Dive

```{r final_model}
final_model <- model_int_scores
```

### Summary

```{r final_summary}
summary(final_model)
```

### Coefficients

```{r coef_table}
coef_df <- as.data.frame(summary(final_model)$coefficients)
coef_df <- round(coef_df, 5)
coef_df$Significant <- ifelse(coef_df[, 4] < 0.05, "***", "")
kable(coef_df[, 1:4] %>% cbind(Significant = coef_df$Significant),
      caption = "Final Model Coefficients")
```

**What the coefficients mean:**

- **Critic_Score**: A higher critic score is associated with higher log(Global Sales). The effect is multiplicative on the actual sales scale since we log-transformed the response.
- **User_Score_100**: The main effect of user score. Once you account for the interaction, the main effect on its own is less straightforward to interpret directly.
- **Critic_Score:User_Score_100**: This is the interesting one. A positive interaction means games that score well with both crtics and users do even better than you'd expect from either score alone, basically there's a bonus for universal acclaim.
- **Year_of_Release**: Sales change linearly over time, picking up the overall industry trend.
- **Genre**: R needs a baseline genre to compare against since you can't include all 12 at once without causing multicollinearity problems. It picks Action because it comes first alphabetically. Every other genre coefficient tells you how much higher or lower log(Sales) is expected compared to Action games.
- **Rating**: Same idea as Genre, E-rated games are the baseline and other ratings are relative to that.

### Interaction Plot

```{r interaction_plot, fig.height=5}
pred_grid <- expand.grid(
  Critic_Score = seq(20, 98, by = 2),
  User_Score_100 = c(40, 70, 90),
  Year_of_Release = median(model_df$Year_of_Release),
  Genre = "Action",
  Rating = "E"
)

pred_grid$predicted <- predict(final_model, newdata = pred_grid)
pred_grid$User_Score_Grp <- factor(pred_grid$User_Score_100,
                                    labels = c("Low (40)", "Mid (70)", "High (90)"))

ggplot(pred_grid, aes(x = Critic_Score, y = predicted, color = User_Score_Grp)) +
  geom_line(linewidth = 1) +
  scale_color_brewer(palette = "Set1") +
  labs(
    title = "How Critic Score Affects Predicted Sales at Different User Score Levels",
    subtitle = "Action games, median year, E rating",
    x = "Critic Score",
    y = "Predicted log(Global Sales)",
    color = "User Score"
  ) +
  theme_minimal(base_size = 12)
```

The lines being non-parallel shows the interaction is real, critic score matters more when user scores are also high.

### Genre Effects

```{r genre_plot, fig.height=5}
genre_coefs <- as.data.frame(summary(final_model)$coefficients)
genre_coefs$term <- rownames(genre_coefs)

genre_main <- genre_coefs %>%
  filter(grepl("^Genre", term)) %>%
  mutate(
    Genre = gsub("Genre", "", term),
    lower = Estimate - 1.96 * `Std. Error`,
    upper = Estimate + 1.96 * `Std. Error`,
    sig = `Pr(>|t|)` < 0.05
  )

ggplot(genre_main, aes(x = Estimate, y = reorder(Genre, Estimate),
                        color = sig, xmin = lower, xmax = upper)) +
  geom_pointrange() +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  scale_color_manual(values = c("FALSE" = "grey70", "TRUE" = "steelblue"),
                     labels = c("not significant", "p < 0.05")) +
  labs(title = "Genre Coefficients Relative to Action",
       x = "Estimate (95% CI)", y = NULL, color = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

### Actual vs Predicted

```{r actual_pred, fig.height=5}
model_df$fitted_final <- fitted(final_model)
model_df$resid_final <- resid(final_model)

ggplot(model_df, aes(x = fitted_final, y = log_Global_Sales)) +
  geom_point(alpha = 0.2, size = 0.7, color = "steelblue") +
  geom_abline(slope = 1, intercept = 0, color = "tomato", linewidth = 1, linetype = "dashed") +
  labs(title = "Actual vs Predicted",
       subtitle = "Points on the dashed line = perfect prediction",
       x = "Predicted", y = "Actual") +
  theme_minimal(base_size = 12)
```

## Diagnostics

```{r diagnostics, fig.width=11, fig.height=8}
par(mfrow = c(2, 2))
plot(final_model, which = 1:4, pch = 20, cex = 0.4,
     col = adjustcolor("steelblue", alpha.f = 0.4))
par(mfrow = c(1, 1))
```

### BP Test

```{r bp_test}
bptest(final_model)
```

The BP test is significant which means there's heteroscedasticity. Even after logging the response there's still some non constant variance, mostly because of extreme blockbuster titles that sit way above the rest of the data.

### Residuals

```{r resid_hist, fig.height=4}
hist(resid(final_model),
     main = "Residuals - Final Model",
     xlab = "Residuals",
     col = "steelblue",
     border = "white",
     breaks = 35)
```

### VIF Check

```{r vif_check}
vif(final_model, type = "predictor") %>%
  as.data.frame() %>%
  tibble::rownames_to_column("Predictor") %>%
  arrange(desc(GVIF)) %>%
  kable(digits = 3, caption = "VIF - Final Model")
```

The main predictors (Critic_Score, User_Score_100, Rating) all have VIF well below 10 which is good. The interaction term will naturally have higher VIF and that's expected.

## Statistical Tests

```{r stat_tests}
tests <- data.frame(
  Test = c(
    "F-test: Full model significance",
    "ANOVA: Full vs Critic x User Score",
    "Breusch-Pagan: Constant variance"
  ),
  Result = c(
    paste0("F = ", round(summary(full_model)$fstatistic[1], 2), ", p < 0.001"),
    paste0("F = ", round(anova(full_model, model_int_scores)$F[2], 2),
           ", p = ", round(anova(full_model, model_int_scores)[2, 6], 6)),
    paste0("BP = ", round(bptest(final_model)$statistic, 2),
           ", p = ", round(bptest(final_model)$p.value, 4))
  )
)

kable(tests, caption = "Key Statistical Tests")
```

## Performance Summary

```{r performance}
perf <- data.frame(
  Metric = c("R-squared", "Adjusted R-squared", "Residual SE", "CV RMSE", "AIC"),
  Value = c(
    round(summary(final_model)$r.squared, 4),
    round(summary(final_model)$adj.r.squared, 4),
    round(summary(final_model)$sigma, 4),
    round(rmse_int_scores, 4),
    round(AIC(final_model), 2)
  )
)

kable(perf, caption = "Final Model Performance")
```

## Alternative Model

If you care more about answering the genre question than raw predictive performance, the Critic x Genre model is worth considering. It scored slightly worse on AIC and CV RMSE but it gives you a clearer picture of how critic scores play out differently across genres.

```{r alt_model}
cat("Final (Critic x User Score) - Adj R2:", round(summary(final_model)$adj.r.squared, 4),
    "| CV RMSE:", round(rmse_int_scores, 4),
    "| AIC:", round(AIC(final_model), 2), "\n")

cat("Alt (Critic x Genre)        - Adj R2:", round(summary(model_int_genre)$adj.r.squared, 4),
    "| CV RMSE:", round(rmse_int_genre, 4),
    "| AIC:", round(AIC(model_int_genre), 2), "\n")
```

## Key Findings

```{r key_coefs}
main_coefs <- as.data.frame(summary(final_model)$coefficients)
main_coefs <- main_coefs[c("Critic_Score", "User_Score_100", "Critic_Score:User_Score_100"), ]
kable(round(main_coefs, 5), caption = "Key Coefficients")
```

- **Critic Score** is a significant positive predictor of global sales. Higher critic scores mean higher expected sales.
- **User Score** has a negative main effect after controlling for critic scores. This sounds counterintuitive but it's because the two scores are correlated, once you control for critic score, user score is picking up a different signal. The real story is in the interaction.
- **The interaction term** is the main finding of this project. Games that are well received by both critics and users sell more than what either score alone would predict. Universal acclaim has a multiplier effect.
- **Genre** matters, some genres have higher expected sales than Action (the baseline) and some have lower. These differences are statistically significant for several genres.
- **ESRB Rating** is also significant. M and E rated games tend to outperform T and E10+ after controlling for everything else.
- The model explains about `r round(summary(final_model)$adj.r.squared * 100, 1)`% of the variance in log(Global Sales) with a CV RMSE of `r round(rmse_int_scores, 4)`. An R-squared around 19% is honestly pretty reasonable for predicting sales, there are a ton of factors that drive commercial success that we don't have data on, like marketing spend and franchise recognition. The F-statistic is highly significant (p < 2.2e-16) which confirms the model is meaningfully better than just guessing the mean.
