---
title: "Video Game Sales Modeling"
author: "Dangelo Feliciano, Anirudh Eswara"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 5,
  dpi = 150
)
```

```{r libs}
library(tidyverse)
library(knitr)
library(lmtest)
library(moments)

source(here::here("R", "01_load_clean.R"))
```

## Data Setup

Before fitting any models we need to transform and filter the data. The big thing here is that Global_Sales is extremely right skewed, so we use log1p() to bring it closer to normal. We also scale User_Score up to 0-100 so it's on the same scale as Critic_Score, which makes interpreting coefficients easier. We drop any rows missing our key predictors since we can't really do anything useful with incomplete observations.

```{r prep}
df <- run_pipeline()

df <- df %>%
  mutate(
    log_Global_Sales = log1p(Global_Sales),
    User_Score_100 = User_Score * 10,
    Genre = as.factor(Genre),
    Rating = if_else(Rating %in% c("E", "E10+", "T", "M"), as.character(Rating), "Other"),
    Rating = as.factor(Rating),
    Platform = as.factor(Platform)
  )

model_df <- df %>%
  filter(
    !is.na(Critic_Score),
    !is.na(User_Score_100),
    !is.na(Rating)
  )

cat("rows for modeling:", nrow(model_df), "\n")
cat("skewness of log_Global_Sales:", round(skewness(model_df$log_Global_Sales), 3), "\n")
```

## Full Model

For our initial model we included everything we thought would reasonably predict global sales. We left out Name, Developer, and Publisher because they have way too many levels and would just overfit. We also left out the regional sales columns (NA_Sales, EU_Sales, etc.) because those literally add up to Global_Sales, so including them would be cheating.

```{r full_model}
full_model <- lm(
  log_Global_Sales ~ Critic_Score + User_Score_100 + Year_of_Release + Genre + Rating,
  data = model_df
)

summary(full_model)
```

```{r full_model_fit}
cat("R-squared:", round(summary(full_model)$r.squared, 4), "\n")
cat("Adj R-squared:", round(summary(full_model)$adj.r.squared, 4), "\n")
cat("Residual SE:", round(summary(full_model)$sigma, 4), "\n")
cat("F-stat p-value:", round(pf(summary(full_model)$fstatistic[1],
                                summary(full_model)$fstatistic[2],
                                summary(full_model)$fstatistic[3],
                                lower.tail = FALSE), 6), "\n")
```

```{r full_model_coefs}
coef_df <- as.data.frame(summary(full_model)$coefficients)
coef_df$Significant <- ifelse(coef_df[, 4] < 0.05, "Yes", "No")
kable(round(coef_df[, 1:4], 5) %>% cbind(Significant = coef_df$Significant),
      caption = "Full Model Coefficients")
```

## Checking Assumptions

```{r resid_vs_fitted, fig.height=4}
plot(fitted(full_model), resid(full_model),
     col = "grey", pch = 20,
     xlab = "Fitted Values", ylab = "Residuals",
     main = "Residuals vs Fitted - Full Model")
abline(h = 0, col = "darkorange", lwd = 2)
```

```{r bp_test}
bptest(full_model)
```

The BP test is significant which means we have heteroscedasticity. This is pretty common with sales data even after the log transform because there are still some extreme outlier titles that are hard to model.

```{r qq_plot}
qqnorm(resid(full_model), main = "Normal Q-Q Plot - Full Model", col = "darkgrey")
qqline(resid(full_model), col = "dodgerblue", lwd = 2)
```

```{r resid_hist, fig.height=4}
hist(resid(full_model),
     main = "Residuals - Full Model",
     xlab = "Residuals",
     col = "darkorange",
     border = "dodgerblue",
     breaks = 30)
```

The Q-Q plot shows some deviation in the tails but overall the residuals look reasonable. The histogram confirms they are roughly bell shaped.

## 10-Fold Cross Validation

We ran 10-fold CV to get a baseline RMSE. Every model we try going forward will be compared to this number.

```{r cv_full}
set.seed(42)

n <- nrow(model_df)
folds <- sample(rep(1:10, length.out = n))
rmse_folds <- numeric(10)

for (k in 1:10) {
  train <- model_df[folds != k, ]
  test <- model_df[folds == k, ]
  fit <- lm(log_Global_Sales ~ Critic_Score + User_Score_100 + Year_of_Release + Genre + Rating,
            data = train)
  preds <- predict(fit, newdata = test)
  rmse_folds[k] <- sqrt(mean((test$log_Global_Sales - preds)^2))
}

cat("Baseline CV RMSE:", round(mean(rmse_folds), 4), "\n")
cat("SD:", round(sd(rmse_folds), 4), "\n")
```

## Trying to Improve the Model

### Stepwise Selection

We tried stepwise selection using AIC to see if any variables should be dropped. It ended up keeping everything, which tells us all the predictors are doing something useful.

```{r stepwise, results='hide'}
step_model <- step(full_model, direction = "both", trace = 0)
```

```{r stepwise_compare}
summary(step_model)
cat("Full model AIC:", round(AIC(full_model), 2), "\n")
cat("Step model AIC:", round(AIC(step_model), 2), "\n")
```

```{r cv_step}
set.seed(42)
rmse_step <- numeric(10)

for (k in 1:10) {
  train <- model_df[folds != k, ]
  test <- model_df[folds == k, ]
  fit <- lm(formula(step_model), data = train)
  preds <- predict(fit, newdata = test)
  rmse_step[k] <- sqrt(mean((test$log_Global_Sales - preds)^2))
}

cat("CV RMSE - Stepwise:", round(mean(rmse_step), 4), "\n")
```

### Critic Score x Genre Interaction

One of our main research questions is whether critic scores matter more in some genres than others. We tried adding a Critic Score x Genre interaction to test this. The model was significant but it didn't beat the score interaction model below in terms of AIC or CV RMSE.

```{r int_genre}
model_int_genre <- lm(
  log_Global_Sales ~ Critic_Score * Genre + User_Score_100 + Year_of_Release + Rating,
  data = model_df
)

summary(model_int_genre)
cat("AIC:", round(AIC(model_int_genre), 2), "\n")
```

```{r cv_int_genre}
set.seed(42)
rmse_int_genre <- numeric(10)

for (k in 1:10) {
  train <- model_df[folds != k, ]
  test <- model_df[folds == k, ]
  fit <- lm(log_Global_Sales ~ Critic_Score * Genre + User_Score_100 + Year_of_Release + Rating,
            data = train)
  preds <- predict(fit, newdata = test)
  rmse_int_genre[k] <- sqrt(mean((test$log_Global_Sales - preds)^2))
}

cat("CV RMSE - Critic x Genre:", round(mean(rmse_int_genre), 4), "\n")
```

### Critic Score x User Score Interaction

We also wanted to test whether games that score well with both critics AND users do better than you would expect from each score alone. This one gave us the best results out of everything we tried.

```{r int_scores}
model_int_scores <- lm(
  log_Global_Sales ~ Critic_Score * User_Score_100 + Year_of_Release + Genre + Rating,
  data = model_df
)

summary(model_int_scores)
cat("AIC:", round(AIC(model_int_scores), 2), "\n")
```

```{r cv_int_scores}
set.seed(42)
rmse_int_scores <- numeric(10)

for (k in 1:10) {
  train <- model_df[folds != k, ]
  test <- model_df[folds == k, ]
  fit <- lm(log_Global_Sales ~ Critic_Score * User_Score_100 + Year_of_Release + Genre + Rating,
            data = train)
  preds <- predict(fit, newdata = test)
  rmse_int_scores[k] <- sqrt(mean((test$log_Global_Sales - preds)^2))
}

cat("CV RMSE - Critic x User Score:", round(mean(rmse_int_scores), 4), "\n")
```

### Polynomial Year

We thought year might have a nonlinear effect since the industry grew quickly then slowed down. A quadratic term improved things over the baseline but not over the score interaction model.

```{r poly_year}
model_poly_year <- lm(
  log_Global_Sales ~ Critic_Score + User_Score_100 + poly(Year_of_Release, 2) + Genre + Rating,
  data = model_df
)

summary(model_poly_year)
cat("AIC:", round(AIC(model_poly_year), 2), "\n")
```

```{r cv_poly_year}
set.seed(42)
rmse_poly_year <- numeric(10)

for (k in 1:10) {
  train <- model_df[folds != k, ]
  test <- model_df[folds == k, ]
  fit <- lm(log_Global_Sales ~ Critic_Score + User_Score_100 + poly(Year_of_Release, 2) + Genre + Rating,
            data = train)
  preds <- predict(fit, newdata = test)
  rmse_poly_year[k] <- sqrt(mean((test$log_Global_Sales - preds)^2))
}

cat("CV RMSE - Poly Year:", round(mean(rmse_poly_year), 4), "\n")
```

### Combined Model

We combined the genre interaction with the polynomial year term to see if putting them together would help. It did better than either alone but still didn't beat the Critic x User Score interaction.

```{r model_best}
model_best <- lm(
  log_Global_Sales ~ Critic_Score * Genre + User_Score_100 + poly(Year_of_Release, 2) + Rating,
  data = model_df
)

summary(model_best)
cat("AIC:", round(AIC(model_best), 2), "\n")
```

```{r cv_best}
set.seed(42)
rmse_best <- numeric(10)

for (k in 1:10) {
  train <- model_df[folds != k, ]
  test <- model_df[folds == k, ]
  fit <- lm(log_Global_Sales ~ Critic_Score * Genre + User_Score_100 + poly(Year_of_Release, 2) + Rating,
            data = train)
  preds <- predict(fit, newdata = test)
  rmse_best[k] <- sqrt(mean((test$log_Global_Sales - preds)^2))
}

cat("CV RMSE - Combined:", round(mean(rmse_best), 4), "\n")
```

## Comparing All Models

```{r comparison}
comparison <- data.frame(
  Model = c(
    "Full Model",
    "Stepwise AIC",
    "Critic x Genre",
    "Critic x User Score (Final)",
    "Polynomial Year",
    "Combined Model"
  ),
  AIC = round(c(
    AIC(full_model), AIC(step_model), AIC(model_int_genre),
    AIC(model_int_scores), AIC(model_poly_year), AIC(model_best)
  ), 2),
  CV_RMSE = round(c(
    mean(rmse_folds), mean(rmse_step), mean(rmse_int_genre),
    mean(rmse_int_scores), mean(rmse_poly_year), mean(rmse_best)
  ), 4),
  Adj_R2 = round(c(
    summary(full_model)$adj.r.squared, summary(step_model)$adj.r.squared,
    summary(model_int_genre)$adj.r.squared, summary(model_int_scores)$adj.r.squared,
    summary(model_poly_year)$adj.r.squared, summary(model_best)$adj.r.squared
  ), 4)
)

kable(comparison, caption = "Model Comparison")
```

The Critic x User Score interaction model came out on top with the lowest AIC (7276) and CV RMSE (0.4098). We're carrying this forward as our final model.

## Diagnostics on the Final Model

```{r final_diag, fig.width=11, fig.height=8}
par(mfrow = c(2, 2))
plot(model_int_scores, which = 1:4, pch = 20, cex = 0.4,
     col = adjustcolor("steelblue", alpha.f = 0.4))
par(mfrow = c(1, 1))
```

```{r final_bp}
bptest(model_int_scores)
```

```{r final_pred, fig.height=5}
model_df$fitted_best <- fitted(model_int_scores)

ggplot(model_df, aes(x = fitted_best, y = log_Global_Sales)) +
  geom_point(alpha = 0.25, size = 0.8, color = "steelblue") +
  geom_abline(slope = 1, intercept = 0, color = "tomato", linewidth = 1, linetype = "dashed") +
  labs(title = "Actual vs Predicted - Final Model",
       subtitle = "Dashed line = perfect fit",
       x = "Predicted log(Global Sales)",
       y = "Actual log(Global Sales)") +
  theme_minimal()
```
