---
title: "Video Game Sales Preprocessing"
author: "Dangelo Feliciano, Anirudh Eswara"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width  = 10,
  fig.height = 5,
  dpi = 150
)
```

## Setup

```{r libs}
library(tidyverse)
library(scales)
library(knitr)

source(here::here("R", "01_load_clean.R"))
```


## 1. Load Raw Data

```{r load}
raw <- load_raw()
glimpse(raw)
```


## 2. Missing Value Overview

```{r missing-table}
col_summary(raw) %>%
  arrange(desc(null_pct)) %>%
  kable(digits = 2, caption = "Missing Values — Raw Dataset")
```

```{r missing-plot, fig.height=4}
col_summary(raw) %>%
  filter(null_pct > 0) %>%
  ggplot(aes(x = reorder(column, null_pct), y = null_pct)) +
  geom_col(fill = "steelblue", color = "white") +
  coord_flip() +
  labs(title = "Missing Value Rate by Column",
       x = NULL, y = "Missing (%)") +
  theme_minimal()
```

> **Observations:**
> - `Critic_Score`, `User_Score`, `Critic_Count`, `User_Count`, `Developer`, and `Rating` all have substantial missingness, these come from Metacritic and were not available for every title.
> - `Year_of_Release` and `Genre` have minor missingness.
> - Sales columns and `Name`/`Platform`/`Publisher` are nearly complete.


## 3. Cleaning Steps

### 3a. Remove Missing / Zero Target

We cannot model rows with no `Global_Sales` value.

```{r clean-target}
df <- raw %>% clean_target()
cat("Rows remaining:", nrow(df), "\n")
```

### 3b. Clean Year of Release

`Year_of_Release` is stored as character (some entries are `"N/A"`). We coerce to integer and drop implausible values outside 1980–2020.

```{r clean-year}
df <- df %>% clean_year()
cat("Year range:", range(df$Year_of_Release), "\n")
```

### 3c. Filter Valid Genres

Drop rows with missing or unrecognised genre values.

```{r clean-genre}
df <- df %>% clean_genre()
cat("Genres present:", paste(sort(unique(df$Genre)), collapse = ", "), "\n")
```

### 3d. Clean User Score

`User_Score` contains `"tbd"` strings where a score hasn't been assigned yet. We replace these with `NA` and convert the column to numeric.

```{r clean-user-score}
df <- df %>% clean_user_score()

cat("User_Score — NA count after cleaning:",
    sum(is.na(df$User_Score)), "\n")
```

### 3e. Standardise ESRB Rating

Non-standard ESRB values are collapsed to `"Other"` and missing ratings are labeled `"Unknown"`.

```{r clean-rating}
df <- df %>% clean_rating()
df %>% count(Rating, sort = TRUE) %>% kable()
```


## 4. Cleaned Dataset Summary

```{r cleaned-summary}
cat("Shape after cleaning:", nrow(df), "rows x", ncol(df), "cols\n\n")

col_summary(df) %>%
  arrange(desc(null_pct)) %>%
  kable(digits = 2, caption = "Missing Values — After Cleaning")
```

> **Remaining missingness** in `Critic_Score`, `User_Score`, `Rating`, etc. reflects games that simply were not reviewed on Metacritic or rated by the ESRB, this is expected and will be addressed at the modeling stage.

---


write_csv(df, here::here("data", "vgsales_clean.csv"))
cat("Saved: data/vgsales_clean.csv\n")

```{r}
df$Developer = as.factor(df$Developer)
df$Publisher = as.factor(df$Publisher)
df$Name = as.factor(df$Name)
df$Platform = as.factor(df$Platform)
df$Genre = as.factor(df$Genre)
```

```{r}
head(df)
str(df)
summary(df)
colSums(is.na(df))
sum(complete.cases(df))
```

```{r}
hist(df$Year_of_Release,
     main = "Year of Release Histogram",
     xlab = "Year",
     ylab = "Frequency")
boxplot(df$Year_of_Release,
     main = "Year of Release Boxplot")
```

```{r}
hist(df$Critic_Score,
     main = "Critic Score Histogram",
     xlab = "Score",
     ylab = "Frequency")
boxplot(df$Critic_Score,
     main = "Critic Score Boxplot")
```

```{r}
hist(df$Critic_Count,
     main = "Critic Count Histogram",
     xlab = "Count",
     ylab = "Frequency")
boxplot(df$Critic_Count,
     main = "Critic Count Boxplot")
```

```{r}
hist(df$User_Score,
     main = "User Score Histogram",
     xlab = "Score",
     ylab = "Frequency")
boxplot(df$User_Score,
     main = "User Score Boxplot")
```


```{r}
hist(df$User_Count,
     main = "User Count Histogram",
     xlab = "Count",
     ylab = "Frequency")
boxplot(df$User_Count,
     main = "User Count Boxplot")
```

```{r}
cor(df[sapply(df, is.numeric)], use = "pairwise.complete.obs")
pairs(df[sapply(df, is.numeric)])
```

```{r}
barplot(table(df$Platform),
     main = "Platform Bar Plot",
     xlab = "Platform",
     ylab = "Frequency", las = 2)
```

```{r}
barplot(table(df$Genre),
     main = "Genre Bar Plot",
     xlab = "Genre",
     ylab = "Frequency", las = 2)
```


```{r}
barplot(table(df$Rating),
     main = "Rating Bar Plot",
     xlab = "Rating",
     ylab = "Frequency", las = 2)
```

```{r}
fit_1 = lm(Global_Sales ~ .-Name-NA_Sales-EU_Sales-JP_Sales-Other_Sales, data = df)
```

```{r}
plot(fitted(fit_1), resid(fit_1), col = "grey", pch = 20,
     xlab = "Fitted", ylab = "Residuals", main = "Data from Model 1")
abline(h = 0, col = "darkorange", lwd = 2)
library(lmtest)
bptest(fit_1)
h = hist(resid(fit_1),
     xlab   = "Residuals",
     main   = "Histogram of Residuals, fit_1",
     col    = "darkorange",
     border = "dodgerblue",
     breaks = 20)
```

```{r}
qqnorm(resid(fit_1), main = "Normal Q-Q Plot, fit_1", col = "darkgrey")
qqline(resid(fit_1), col = "dodgerblue", lwd = 2)
```