---
title: "Video Game Sales Preprocessing"
author: "Dangelo Feliciano, Anirudh Eswara"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 5,
  dpi = 150
)
```

```{r libs}
library(tidyverse)
library(scales)
library(knitr)

source(here::here("R", "01_load_clean.R"))
```

## Loading the Data

```{r load}
raw <- load_raw()
glimpse(raw)
```

## Missing Values

The first thing we wanted to look at was how much data is actually missing. A lot of the review-related columns come from Metacritic which didn't cover every game, especially older titles, so we expected some gaps there.

```{r missing_table}
col_summary(raw) %>%
  arrange(desc(null_pct)) %>%
  kable(digits = 2, caption = "Missing Values in Raw Dataset")
```

```{r missing_plot, fig.height=4}
col_summary(raw) %>%
  filter(null_pct > 0) %>%
  ggplot(aes(x = reorder(column, null_pct), y = null_pct)) +
  geom_col(fill = "steelblue", color = "white") +
  coord_flip() +
  labs(title = "Missing Value Rate by Column",
       x = NULL, y = "Missing (%)") +
  theme_minimal()
```

As expected, Critic_Score, User_Score, Critic_Count, User_Count, Developer, and Rating all have a lot of missing data since those come from Metacritic and weren't available for every title. Year_of_Release and Genre have minor missingness and the sales columns are almost complete.

## Cleaning

### Drop Rows with No Sales

We can't model games with no Global_Sales value so we remove those first.

```{r clean_target}
df <- raw %>% clean_target()
cat("rows remaining:", nrow(df), "\n")
```

### Fix Year of Release

Year_of_Release came in as a character column with some entries saying "N/A". We convert it to an integer and drop anything outside 1980-2020 since those are either errors or too far out of range to be useful.

```{r clean_year}
df <- df %>% clean_year()
cat("year range:", range(df$Year_of_Release), "\n")
```

### Filter Genres

A small number of rows had missing or unrecognized genre values. We dropped those.

```{r clean_genre}
df <- df %>% clean_genre()
cat("genres:", paste(sort(unique(df$Genre)), collapse = ", "), "\n")
```

### Fix User Score

User_Score had "tbd" entries for games that hadn't been scored yet. We replaced those with NA and converted the column to numeric.

```{r clean_user_score}
df <- df %>% clean_user_score()
cat("User_Score NAs after cleaning:", sum(is.na(df$User_Score)), "\n")
```

### Clean Up ESRB Ratings

There were a handful of non-standard rating values. We collapsed those into "Other" and filled missing ratings as "Unknown".

```{r clean_rating}
df <- df %>% clean_rating()
df %>% count(Rating, sort = TRUE) %>% kable()
```

## Where Things Stand After Cleaning

```{r cleaned_summary}
cat("shape:", nrow(df), "rows x", ncol(df), "cols\n\n")

col_summary(df) %>%
  arrange(desc(null_pct)) %>%
  kable(digits = 2, caption = "Missing Values After Cleaning")
```

The remaining missingness in Critic_Score, User_Score, Rating, and Developer is expected - these are just games that weren't on Metacritic or didn't get an ESRB rating. We'll handle those at the modeling stage by only using complete cases.

## Convert to Factors and Explore

```{r factors}
df$Developer <- as.factor(df$Developer)
df$Publisher <- as.factor(df$Publisher)
df$Name <- as.factor(df$Name)
df$Platform <- as.factor(df$Platform)
df$Genre <- as.factor(df$Genre)
```

```{r explore}
head(df)
str(df)
summary(df)
colSums(is.na(df))
sum(complete.cases(df))
```

## Distributions

### Year of Release

```{r year_dist}
hist(df$Year_of_Release,
     main = "Year of Release",
     xlab = "Year",
     ylab = "Frequency")
boxplot(df$Year_of_Release,
        main = "Year of Release Boxplot")
```

### Critic Score

```{r critic_dist}
hist(df$Critic_Score,
     main = "Critic Score",
     xlab = "Score",
     ylab = "Frequency")
boxplot(df$Critic_Score,
        main = "Critic Score Boxplot")
```

### Critic Count

```{r critic_count_dist}
hist(df$Critic_Count,
     main = "Critic Count",
     xlab = "Count",
     ylab = "Frequency")
boxplot(df$Critic_Count,
        main = "Critic Count Boxplot")
```

### User Score

```{r user_dist}
hist(df$User_Score,
     main = "User Score",
     xlab = "Score",
     ylab = "Frequency")
boxplot(df$User_Score,
        main = "User Score Boxplot")
```

### User Count

```{r user_count_dist}
hist(df$User_Count,
     main = "User Count",
     xlab = "Count",
     ylab = "Frequency")
boxplot(df$User_Count,
        main = "User Count Boxplot")
```

## Correlations

```{r correlations}
cor(df[sapply(df, is.numeric)], use = "pairwise.complete.obs")
pairs(df[sapply(df, is.numeric)])
```

## Categorical Variables

```{r platform_bar}
barplot(table(df$Platform),
        main = "Games by Platform",
        xlab = "Platform",
        ylab = "Count", las = 2)
```

```{r genre_bar}
barplot(table(df$Genre),
        main = "Games by Genre",
        xlab = "Genre",
        ylab = "Count", las = 2)
```

```{r rating_bar}
barplot(table(df$Rating),
        main = "Games by Rating",
        xlab = "Rating",
        ylab = "Count", las = 2)
```

## Initial Model

Before getting into the full modeling file we fit a quick initial model just to get a feel for the data. We used raw Global_Sales as the response here which we knew would be skewed - the diagnostics below show that clearly. The log transformation is applied in the modeling file.

```{r fit_1}
fit_1 <- lm(Global_Sales ~ . -Name -NA_Sales -EU_Sales -JP_Sales -Other_Sales, data = df)
```

```{r fit_1_diagnostics}
plot(fitted(fit_1), resid(fit_1), col = "grey", pch = 20,
     xlab = "Fitted", ylab = "Residuals", main = "Residuals vs Fitted - fit_1")
abline(h = 0, col = "darkorange", lwd = 2)

library(lmtest)
bptest(fit_1)

hist(resid(fit_1),
     xlab = "Residuals",
     main = "Histogram of Residuals - fit_1",
     col = "darkorange",
     border = "dodgerblue",
     breaks = 20)
```

```{r fit_1_qq}
qqnorm(resid(fit_1), main = "Normal Q-Q Plot - fit_1", col = "darkgrey")
qqline(resid(fit_1), col = "dodgerblue", lwd = 2)
```

The residuals are clearly not normal and the BP test confirms heteroscedasticity. This is why we log-transform Global_Sales before fitting the full model.

## Save Clean Data

```{r save}
write_csv(df, here::here("data", "vgsales_clean.csv"))
cat("saved vgsales_clean.csv\n")
```
